---
title: "Reproducible Research: Peer Assessment 1"
author: "Mark Nemeth"
date: "7/9/2021"
output: html_document
#md_document:
#html_document:
#keep_md:true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction / Purpose

It is now possible to collect a large amount of data about personal movement using activity monitoring devices such as a Fitbit, Nike Fuelband, or Jawbone Up. These type of devices are part of the "quantified self" movement -- a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. But these data remain under-utilized both because the raw data are hard to obtain and there is a lack of statistical methods and software for processing and interpreting the data.

This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.


## Loading and preprocessing the data

```{r load_data, echo=TRUE}

library(dplyr)
library(ggplot2)

classes<-c("numeric","character","numeric")
activity_df<-read.csv("./data/activity.csv", colClasses=classes, header=TRUE)

activity_df$date<-as.Date(activity_df$date, "%Y-%m-%d")
str(activity_df)
head(activity_df)
summary(activity_df)

```

We see the data is tidy: It contains single observation records for a particular anonymous individual, where each record records the number of steps measured in an indexed time interval of the day, for each date.  The records are number of steps (numeric), date (calendar date - character read), time interval of day (numeric).

- steps: Number of steps taking in a 5-minute interval (missing values are coded as **NA**  
- date: The date on which the measurement was taken in YYYY-MM-DD format  
- interval: Identifier for the 5-minute interval in which measurement was taken  


## What is mean total number of steps taken per day?
We note that there are 2,304 NA's in the data file.  Our instructions are that we can ignore missings in this part of our work, so we remove them from our daily build.

```{r dailysteps}

dailySteps <- aggregate(activity_df$steps[!is.na(activity_df$steps)], by=list(activity_df$date[!is.na(activity_df$steps)]), FUN=sum)

names(dailySteps)[1]<-"date"
names(dailySteps)[2]<-"steps"

ggplot(data=dailySteps, aes(x=date, y=steps, col="blue")) + geom_bar(stat="identity") +
        scale_x_date(breaks="1 day",
                     limits=c(as.Date("2012-10-01"),as.Date("2012-11-30"))) +
        ylab("Steps") + xlab("Date") + 
        theme_bw() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

summary(dailySteps$steps)

```

And, so we find also that the daily **mean** number of steps is **10,766** and the **median** is **10,765**.

## What is the average daily activity pattern?



## Imputing missing values


## Are there differences in activity patterns between weekdays and weekends?


